---
title: "Find nearest genes and perform overrepresentation (pathway) analysis"
output: html_document
---

```{r}
library(ggplot2)
library(tidyverse)
#set_here("/Users/mary/non_dropbox/exps/exp059_snp_to_sc/")
setwd("/Users/mary/non_dropbox/exps/exp059_snp_to_sc/")
library(here)
library(ChIPseeker)

# Loads cytoimmgen CHEERS results, including the snps overlapping peaks 
load(here("res/cheers_cytoimmgen_processed.RData"))

# T cells
t_celllist <- c('naive_16H_UNS', 'naive_D5_UNS','memory_16H_UNS', 'memory_D5_UNS',  'naive_16H_TH0', 'naive_16H_TH1', 'naive_16H_TH2', 'naive_16H_TH17', 'naive_16H_ITREG', 'naive_16H_IL10', 'naive_16H_IL21', 'naive_16H_IL27', 'naive_16H_IFNB', 'naive_16H_TNFA', 'naive_D5_TH0',  'naive_D5_TH1', 'naive_D5_TH2', 'naive_D5_TH17', 'naive_D5_ITREG', 'naive_D5_IL10', 'naive_D5_IL21', 'naive_D5_IL27', 'naive_D5_IFNB', 'naive_D5_TNFA', 'memory_16H_TH0', 'memory_16H_TH1', 'memory_16H_TH2', 'memory_16H_TH17', 'memory_16H_ITREG', 'memory_16H_IL10', 'memory_16H_IL21', 'memory_16H_IL27', 'memory_16H_IFNB', 'memory_16H_TNFA', 'memory_D5_TH0', 'memory_D5_TH1', 'memory_D5_TH2', 'memory_D5_TH17', 'memory_D5_ITREG', 'memory_D5_IL10', 'memory_D5_IL21', 'memory_D5_IL27', 'memory_D5_IFNB', 'memory_D5_TNFA')

lynall_get_top_peaks <- function(pValsFile=NULL, uniquePeaksFile=NULL, celllist=NULL){
  # Function gets top 10% of peaks active in the significantly enriched cell subsets (Bonferroni p<0.05)
  # Output is vector of peaks
  pvals <- read.table(pValsFile, sep="\t")
  library(magrittr)
  pvals %<>% dplyr::rename(cell_condition=V1,value=V2)
  print(sprintf("There are %d used cell types as follows:", length(celllist))) 
  print(celllist)
  if (length(celllist)==0){
    return(NA)
  } else {
    peaks <- read.table(uniquePeaksFile, sep="\t", header=T)[,c("chr","start","end",celllist)]
    peaks$peak <- paste(peaks$chr, peaks$start, peaks$end, sep = '_')
    # For each cell type, convert to percetile ranks (scales 0 --> 1)
    peaks_rank <- dplyr::mutate(peaks, across(.cols=all_of(celllist), percent_rank))
    # Keep any peaks with percentile rank >=0.9 in any of the sig cell types
    peaks_top <- peaks_rank %>% filter_at(vars(all_of(celllist)), .vars_predicate = any_vars(. >= 0.9)) %>% dplyr::arrange(peak)
  }
}

top_peaks <- list()
# Include those disorders showing T cell enrichment in cytoimmgen
for (i in c("bmi","mdd","crossdis","szs","ra")){
  print(i)
  top_peaks[[i]] <- lynall_get_top_peaks(pValsFile=here(paste0("res/cheersout_500/",i,"_cytoimmgen_H3K27ac_disease_enrichment_pValues.txt")), uniquePeaksFile=here(paste0("res/cheersout_500/",i,"_cytoimmgen_H3K27ac_uniquePeaks.txt")), celllist=t_celllist)
}

length(top_peaks)

```

Prepare peaks as granges for ChIPseeker
```{r}

library(GenomicRanges)
peaks_top_granges <- lapply(top_peaks, function(x) {GenomicRanges::makeGRangesFromDataFrame(x,
                         keep.extra.columns=FALSE,
                         ignore.strand=TRUE,
                         seqnames.field="chr",
                         start.field="start",
                         end.field="end",
                         starts.in.df.are.0based=FALSE)})

```

Load organism databases required for seq2gene
```{r}

library(org.Hs.eg.db)
library(EnsDb.Hsapiens.v86) 
edb <- EnsDb.Hsapiens.v86
genomeStyles()$Homo_sapiens
seqlevelsStyle(edb) <- "UCSC"

```

Get union of exons / introns overlapping peaks and nearest genes. Main modification of lynall_seq2gene compared to CHiPseeker::seq2gene is to retain information about what the peak is overlapping in the function output. A single peak can overlap multiple genomic features (an exon, an intron, a promoter of another gene etc.)
```{r}

source(here("workflow/scripts/functions/lynall_seq2gene.R"))
out <- lapply(peaks_top_granges, function(x) {lynall_seq2gene(x, tssRegion = c(-3000, 3000), flankDistance = 10000, TxDb=edb, sameStrand = FALSE)})

```

Over-representation analysis using pooled reactome and GO (Biological Processes) pathways.

```{r}

# Make term2gene for reactome
reactome <- read.csv("/Users/mary/non_dropbox/exps/exp042_meninges_stress_dropseq/data/raw/reactome_download/Ensembl2Reactome_All_Levels.txt", sep="\t", header=F)
react <- reactome %>% dplyr::filter(grepl("R-HSA-",V2)) %>% dplyr::filter(grepl("ENSG",V1)) 
human_t2g_reactome <- data.frame(gs_name=react$V4, ens=react$V1)

# Get GO term2gene
#library(remotes)
#remotes::install_version("msigdbr","7.4.1")
library(msigdbr) # Requires version 7.4.1
tmp = msigdbr(species = "human", category = "C5", subcategory = "GO:BP")
human_t2g_go <- data.frame(gs_name=tmp$gs_name, ens=tmp$ensembl_gene, symbol=tmp$human_gene_symbol)

# Prettify
human_t2g_go$gs_name <- gsub("GOBP_","",human_t2g_go$gs_name)
human_t2g_go$gs_name <- gsub("_"," ",human_t2g_go$gs_name)
human_t2g_go$gs_name <- paste0(human_t2g_go$gs_name," (GO)")

library(magrittr)
human_t2g_reactome$gs_name %<>% toupper
human_t2g_reactome$gs_name <- paste0(human_t2g_reactome$gs_name," (Reactome)")

human_t2g_pooled <- bind_rows(human_t2g_go, human_t2g_reactome)
sample_n(human_t2g_pooled,20)

# Do pooled reactome and GO over-representation analysis
#remotes::install_url("https://bioconductor.org/packages/3.14/bioc/src/contrib/clusterProfiler_4.2.2.tar.gz")
# NB. enricher uses phyper(...,lower.tail = FALSE) i.e. one-sided hypergeometric test
library(clusterProfiler)
enrichout_pooled <- lapply(out, function(x){enricher(gene=x$unique_genes,minGSSize = 10, maxGSSize = 500, TERM2GENE = human_t2g_pooled)})

# And show the sig results
enrichout_pooled_readable <- lapply(enrichout_pooled, function(x) {setReadable(x, 'org.Hs.eg.db', 'ENSEMBL')})
result_pooled <- lapply(enrichout_pooled_readable, function(x){x@result[x@result$p.adjust<0.05,c("p.adjust", "geneID","Count")]})

```

Print results to file

Unique gene lists for each disorder
```{r}

nearestgenes <- lapply(out, function(x){x$full_dataframe %>% select(peak_location, gene, symbol, overlap_type)})

uniquenearestgenes <- lapply(nearestgenes, function(x){x %>% select(gene,symbol) %>% distinct()})

# FOR SUPPLEMENTARY TABLE 5
for (i in c("bmi","mdd","crossdis","szs", "ra")){
  write.table(uniquenearestgenes[[i]], file=here(paste0("res/overrepresentation_tcells_uniquenearestgenes_",i,".tsv")), row.names = F, quote=F, sep="\t")
}

# Number of unique nearest genes
lapply(uniquenearestgenes,nrow)
#bmi 592
#mdd 65
#crossdis 99
#szs 100
#ra 53

# The unique nearest genes found in significantly enriched pathways 
uniquegenesinpathway <- lapply(enrichout_pooled_readable, function(x){unique(strsplit(paste(x[,"geneID"],collapse="/"),"/")[[1]])}) 

# Proportion in enriched pathways
for (i in c("bmi","mdd","crossdis","szs", "ra")){
  print(i)
  print(sprintf("Total input genes: %d", nrow(uniquenearestgenes[[i]])))
  print(sprintf("Total genes in significantly enriched pathways: %d", length(uniquegenesinpathway[[i]])))
  print(sprintf("Percent in pathway = %f", length(uniquegenesinpathway[[i]])/nrow(uniquenearestgenes[[i]])*100))
}

# [1] "bmi"
# [1] "Total input genes: 592"
# [1] "Total genes in significantly enriched pathways: 0"
# [1] "Percent in pathway = 0.000000"
# [1] "mdd"
# [1] "Total input genes: 65"
# [1] "Total genes in significantly enriched pathways: 8"
# [1] "Percent in pathway = 12.307692"
# [1] "crossdis"
# [1] "Total input genes: 99"
# [1] "Total genes in significantly enriched pathways: 21"
# [1] "Percent in pathway = 21.212121"
# [1] "szs"
# [1] "Total input genes: 100"
# [1] "Total genes in significantly enriched pathways: 32"
# [1] "Percent in pathway = 32.000000"
# [1] "ra"
# [1] "Total input genes: 53"
# [1] "Total genes in significantly enriched pathways: 23"
# [1] "Percent in pathway = 43.396226"

```

The below are the easiest way to see the genes driving the enrichment results - have transferred these onto the figure S8
```{r}
for (i in c("bmi","mdd","crossdis","szs", "ra")){
  write.table(result_pooled[[i]], file=here(paste0("res/overrepresentation_tcells_",i,".tsv")), row.names = T, quote=F, sep="\t")
}

for (i in c("bmi","mdd","crossdis","szs", "ra")){
  write.table(out[[i]]$unique_genes, file=here(paste0("res/overrepresentation_tcells_uniquegenes_",i,".tsv")), row.names = F, col.names=T, quote=F, sep="\t")
}

# Proportion of genes picked out in overrepresentation analysis

```

Make GSEA plots for supplementary, showing top 10 significant pathways
Show -log10 p-value on x-axis and colour by gene ratio, filtering on p(adj)<0.05
```{r}

# What is the range of geneRatios across all?
grmax <- NULL
grmin <- NULL
for (i in 1:length(enrichout_pooled_readable)){
  gr <- enrichout_pooled_readable[[i]]@result %>% dplyr::filter(p.adjust<0.05) 
  gr <- gr %>% rowwise() %>% dplyr::mutate(geneRatio = eval(parse(text = GeneRatio)))
  grmax[i] <- max(gr$geneRatio)
  grmin[i] <- min(gr$geneRatio)
}
grmax
grmin
# Choose visualization range 0 --> 0.4 as this will cover the range for all disorders

lynall_ora_plot <- function(oraout, name="Unnamed", padjthh=0.05, picwidth=15, topn=10){
  df <- oraout@result %>% dplyr::arrange(p.adjust)
  df <- df[1:topn,]
  df %<>% dplyr::filter(p.adjust<padjthh) 
  df$ID %<>% factor(levels=rev(unique(df$ID)), ordered = T)
  df$geneRatio <- sapply(df$GeneRatio, function(x){eval(parse(text = x))})
  p <- ggplot(df, aes(x = -log10(pvalue), y = ID, color=geneRatio)) + 
        geom_point(size=4) + 
        theme_bw(base_size = 14) +
        scale_color_gradient(low="yellow",high="red", limits=c(0,0.32), breaks=c(0,0.1,0.2,0.3)) +
        ylab(NULL) +
        expand_limits(x=0) +
        theme(legend.title = element_blank()) +
        xlab("-log10(p-value)") +
        ggtitle(name)
  print(p)
  ggsave(p,path=here::here("pics/"),filename=paste0("ora_out_",name,".pdf"),width=picwidth,height=(2+nrow(df)*0.11))
  p
}

# Plot results for T cell enriched disorders i.e. Xdiss, SZS, MDD, RA, BMI. Nil sig for BMI so not shown
plots <- list()
for (i in c("bmi","mdd","crossdis","szs", "ra")){
   plots[[i]] <- lynall_ora_plot(oraout=enrichout_pooled_readable[[i]], name=i, padjthh=0.05, picwidth=10, topn=10)
   print(plots[[i]])
}

g <- cowplot::plot_grid(plotlist=list(plots$crossdis, plots$szs, plots$mdd, plots$ra), ncol=1, align="v", rel_heights=c(1,1,0.45,1))

ggsave(g, filename=here("pics/ora_for_supp_select.pdf"), width=11, height=12)
```


