---
title: "Process and visualise Soskic dataset (cytoimmgen) CHEERS results"
output: html_document
---

```{r}

library(ggplot2)
library(tidyverse)
#set_here("/Users/mary/non_dropbox/exps/exp059_snp_to_sc/")
setwd("/Users/mary/non_dropbox/exps/exp059_snp_to_sc/")
library(here)
```

Load cytoimmgen enhancer results
```{r}

cytoimmgen_H3K27ac_p <- list()
cytoimmgen_H3K27ac_meanrank <- list()

# XDISS
cytoimmgen_H3K27ac_p[["Crossdisorder psych"]] <- read.table(here("res/cheersout_500/crossdis_cytoimmgen_H3K27ac_disease_enrichment_pValues.txt"), sep="\t")
cytoimmgen_H3K27ac_meanrank[["Crossdisorder_psych"]] <- read.table(here("res/cheersout_500/crossdis_cytoimmgen_H3K27ac_disease_enrichment_observedMeanRank.txt"), sep="\t")

# SZS
cytoimmgen_H3K27ac_p[["Schizophrenia"]] <- read.table(here("res/cheersout_500/szs_cytoimmgen_H3K27ac_disease_enrichment_pValues.txt"), sep="\t")
cytoimmgen_H3K27ac_meanrank[["Schizophrenia"]] <- read.table(here("res/cheersout_500/szs_cytoimmgen_H3K27ac_disease_enrichment_observedMeanRank.txt"), sep="\t")

# ALZ
cytoimmgen_H3K27ac_p[["Alzheimer's Disease"]] <- read.table(here("res/cheersout_500/alz_cytoimmgen_H3K27ac_disease_enrichment_pValues.txt"), sep="\t")
cytoimmgen_H3K27ac_meanrank[["Alzheimer's Disease"]] <- read.table(here("res/cheersout_500/alz_cytoimmgen_H3K27ac_disease_enrichment_observedMeanRank.txt"), sep="\t")

# BMI
cytoimmgen_H3K27ac_p[["BMI"]] <- read.table(here("res/cheersout_500/bmi_cytoimmgen_H3K27ac_disease_enrichment_pValues.txt"), sep="\t")
cytoimmgen_H3K27ac_meanrank[["BMI"]] <- read.table(here("res/cheersout_500/bmi_cytoimmgen_H3K27ac_disease_enrichment_observedMeanRank.txt"), sep="\t")

# ADHD
cytoimmgen_H3K27ac_p[["ADHD"]] <- read.table(here("res/cheersout_500/adhd_cytoimmgen_H3K27ac_disease_enrichment_pValues.txt"), sep="\t")
cytoimmgen_H3K27ac_meanrank[["ADHD"]] <- read.table(here("res/cheersout_500/adhd_cytoimmgen_H3K27ac_disease_enrichment_observedMeanRank.txt"), sep="\t")

# MDD
cytoimmgen_H3K27ac_p[["Depression"]] <- read.table(here("res/cheersout_500/mdd_cytoimmgen_H3K27ac_disease_enrichment_pValues.txt"), sep="\t")
cytoimmgen_H3K27ac_meanrank[["Depression"]] <- read.table(here("res/cheersout_500/mdd_cytoimmgen_H3K27ac_disease_enrichment_observedMeanRank.txt"), sep="\t")

# BPD
cytoimmgen_H3K27ac_p[["Bipolar Disorder"]] <- read.table(here("res/cheersout_500/bip_cytoimmgen_H3K27ac_disease_enrichment_pValues.txt"), sep="\t")
cytoimmgen_H3K27ac_meanrank[["Bipolar Disorder"]] <- read.table(here("res/cheersout_500/bip_cytoimmgen_H3K27ac_disease_enrichment_observedMeanRank.txt"), sep="\t")

# RA
cytoimmgen_H3K27ac_p[["Rheumatoid arthritis"]] <- read.table(here("res/cheersout_500/ra_cytoimmgen_H3K27ac_disease_enrichment_pValues.txt"), sep="\t")
cytoimmgen_H3K27ac_meanrank[["Rheumatoid arthritis"]] <- read.table(here("res/cheersout_500/ra_cytoimmgen_H3K27ac_disease_enrichment_observedMeanRank.txt"), sep="\t")


```



```{r}

library(purrr)
library(magrittr)
chm <- purrr::map_df(cytoimmgen_H3K27ac_p, ~as.data.frame(.x), .id="condition")
chm %<>% dplyr::rename(cell_condition=V1,value=V2)
chm$condition <- factor(chm$condition , levels = c("Crossdisorder psych","Schizophrenia","Bipolar Disorder", "Depression", "ADHD", "BMI", "Alzheimer's Disease", "Rheumatoid arthritis"))
# Note no ASD as insufficient hits
head(chm)

df <- read.table(here("res/cheersout_500/cytoimmgen_celltypes.txt"), header=F)
colnames(df) <- "cell_condition"
df %<>% separate(cell_condition, into=c("category","time","cell_type"), sep="_", remove=F)
head(df)

cytoimmgen_ncomparisons <- dim(df)[1]
print(cytoimmgen_ncomparisons)

# Rename the timepoints
df$cell_time <- ifelse(df$time=="6H", "macrophage_6H", ifelse(
  df$time=="24H", "macrophage_24H", ifelse(
    df$time=="D5" & df$category=="memory", "memory_D5", ifelse(
      df$time=="D5" & df$category=="naive", "naive_D5", ifelse(
        df$time=="16H" & df$category=="memory","memory_16H", ifelse(
          df$time=="16H" & df$category=="naive","naive_16H", NA))))))

chm_melt <- left_join(chm, df, by="cell_condition")
head(chm_melt)

```


```{r}

cytoimmgen_celllist <- rev(c('naive_16H_UNS', 'naive_D5_UNS','memory_16H_UNS', 'memory_D5_UNS',  'naive_16H_TH0', 'naive_16H_TH1', 'naive_16H_TH2', 'naive_16H_TH17', 'naive_16H_ITREG', 'naive_16H_IL10', 'naive_16H_IL21', 'naive_16H_IL27', 'naive_16H_IFNB', 'naive_16H_TNFA', 'naive_D5_TH0',  'naive_D5_TH1', 'naive_D5_TH2', 'naive_D5_TH17', 'naive_D5_ITREG', 'naive_D5_IL10', 'naive_D5_IL21', 'naive_D5_IL27', 'naive_D5_IFNB', 'naive_D5_TNFA', 'memory_16H_TH0', 'memory_16H_TH1', 'memory_16H_TH2', 'memory_16H_TH17', 'memory_16H_ITREG', 'memory_16H_IL10', 'memory_16H_IL21', 'memory_16H_IL27', 'memory_16H_IFNB', 'memory_16H_TNFA', 'memory_D5_TH0', 'memory_D5_TH1', 'memory_D5_TH2', 'memory_D5_TH17', 'memory_D5_ITREG', 'memory_D5_IL10', 'memory_D5_IL21', 'memory_D5_IL27', 'memory_D5_IFNB', 'memory_D5_TNFA',  'macrophage_6H_UNS', 'macrophage_24H_UNS', 'macrophage_6H_IFNG', 'macrophage_6H_IL4', 'macrophage_6H_IL26', 'macrophage_6H_TNFA', 'macrophage_24H_IFNG', 'macrophage_24H_IL4', 'macrophage_24H_IL23', 'macrophage_24H_IL26', 'macrophage_24H_TNFA'))

timepointlabeller = c("macrophage_24H"="#A6761D","macrophage_6H" = "#E6AB02", "memory_D5" = "#7570B3", 'memory_16H' = "#E7298A", "naive_D5" =  "#1B9E77", "naive_16H" =  "#D95F02")

```

```{r}
ncomp=cytoimmgen_ncomparisons

chm_melt$value[chm_melt$value == 0] <- 1e-16

# Order for graph
chm_melt$cell_condition <- factor(chm_melt$cell_condition, levels=cytoimmgen_celllist)
chm_melt$cell_time <- factor(chm_melt$cell_time, levels=c("macrophage_24H", "macrophage_6H","memory_D5",'memory_16H',"naive_D5","naive_16H"))

library(viridis)
chm_melt$Significance <- factor(ifelse(chm_melt$value<(0.05/ncomp), "Bonf P<0.05", ifelse(chm_melt$value<0.05, "raw P<0.05","Not sig")), levels=c("Not sig","raw P<0.05","Bonf P<0.05"), ordered=T)
                               
# And add p-value rank (ranks smallest to largest)
chm_melt <- chm_melt %>% group_by(condition) %>% mutate(immune_rank=rank(value))

# Print top results
x <- as.data.frame(chm_melt %>% arrange(condition, value) %>% filter(value<0.05) %>% select(condition, cell_condition, value, category, time, cell_type, cell_time))
print(x)
write.table(x, quote=F, row.names = F, sep="\t", file = here(paste0("res/CHEERS_cytoimmgen_allconditions_significant.csv")))

```

FIGURES

```{r}
# Close-up on Xdiss, SZS, MDD, BMI, AD 
p <- ggplot(chm_melt[chm_melt$condition %in% c("Crossdisorder psych","Schizophrenia","Depression","BMI","Alzheimer's Disease"),], aes(cell_condition, -log10(value))) + 
  geom_bar(aes(fill = cell_time), position = "dodge", stat="identity")  +
  facet_wrap(~condition, scale="free_x", ncol = 5) +
  ylab(bquote(~'-log' [10]~ 'p-value')) + 
  xlab("") +
  scale_x_discrete(breaks = chm_melt$cell_condition, labels=chm_melt$cell_type) + 
  geom_hline(yintercept = -log10(0.05), linetype="dotted", color = 'black') + 
  geom_hline(yintercept = -log10(0.05/ncomp), color = 'black') +
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 15), 
        panel.grid = element_blank(), 
        axis.ticks = element_blank(),
        axis.text.y  = element_text(size=10),
        panel.background = element_rect(color="light grey",fill=NA),
        legend.background = element_blank())+ 
  coord_flip() + 
  scale_fill_manual("Cell group", values=timepointlabeller, guide = guide_legend(reverse = TRUE)) 
p
ggsave(p, filename = "CHEERS_cytoimmgen_enhancers_mini_forAI.pdf", path=here("pics"), width=9.7, height=7)

# FIGURE: version for supplement showing all disorders. 
p <- ggplot(chm_melt, aes(cell_condition, -log10(value))) + 
  geom_bar(aes(fill = cell_time), position = "dodge", stat="identity")  +
  facet_wrap(~condition, scale="free_x", ncol = 4) +
  ylab(bquote(~'-log' [10]~ 'p-value')) + 
  xlab("") +
  scale_x_discrete(breaks = chm_melt$cell_condition, labels=chm_melt$cell_type) + 
  geom_hline(yintercept = -log10(0.05), linetype="dotted", color = 'black') + 
  geom_hline(yintercept = -log10(0.05/ncomp), color = 'black') +
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 15), 
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y  = element_text(size=10),
        panel.background = element_rect(color="grey",fill=NA),
        legend.background = element_blank())+ 
  coord_flip() + 
  scale_fill_manual("Cell group", values=timepointlabeller, guide = guide_legend(reverse = TRUE)) 
p
ggsave(p, filename = "CHEERS_cytoimmgen_enhancers_all_forAI.pdf", path=here("pics"), width=10, height=14)

```

Prepare data to plot heatmap of individual peaks
```{r}

anno <- df
rownames(anno) <- anno$cell_condition
anno <- data.frame(anno[,c(4,5)])
head(anno)
annoCols = list(cell_time = timepointlabeller)

# Mean from uniform distribution Mean is 1/2 (a + b) where a and b are minimum and max; number of peaks is 127723
mean_cytoimmgen = (1+127723)/2

get_peaks <- function(uniquePeaksFile=NULL, mean=NULL, celllist=NULL){
  peaks <- read.table(uniquePeaksFile, sep="\t", header=T)
  row.names(peaks) <- paste(peaks$chr, peaks$start, peaks$end, sep = '_')
  peaks <- peaks[c("chr", "start", "end", celllist)]
  peaks <- peaks[,4:dim(peaks)[2]]
  peaks[1:4,1:4]
  #Normalize scores by mean
  peaks <- peaks/mean
}

# Get peaks
peaks=list()
peaks$Depression <- get_peaks(here("res/cheersout_500/mdd_cytoimmgen_H3K27ac_uniquePeaks.txt"), mean=mean_cytoimmgen, celllist=cytoimmgen_celllist)
peaks$`Crossdisorder psych` <- get_peaks(here("res/cheersout_500/crossdis_cytoimmgen_H3K27ac_uniquePeaks.txt"), mean=mean_cytoimmgen, celllist=cytoimmgen_celllist)
peaks$Schizophrenia <- get_peaks(here("res/cheersout_500/szs_cytoimmgen_H3K27ac_uniquePeaks.txt"), mean=mean_cytoimmgen, celllist=cytoimmgen_celllist)
peaks$`Bipolar Disorder` <- get_peaks(here("res/cheersout_500/bip_cytoimmgen_H3K27ac_uniquePeaks.txt"), mean=mean_cytoimmgen, celllist=cytoimmgen_celllist)
peaks$ADHD <- get_peaks(here("res/cheersout_500/adhd_cytoimmgen_H3K27ac_uniquePeaks.txt"), mean=mean_cytoimmgen, celllist=cytoimmgen_celllist)
peaks$BMI <- get_peaks(here("res/cheersout_500/bmi_cytoimmgen_H3K27ac_uniquePeaks.txt"), mean=mean_cytoimmgen, celllist=cytoimmgen_celllist)
peaks$`Alzheimer's Disease` <- get_peaks(here("res/cheersout_500/alz_cytoimmgen_H3K27ac_uniquePeaks.txt"), mean=mean_cytoimmgen, celllist=cytoimmgen_celllist)
peaks$`Rheumatoid arthritis` <- get_peaks(here("res/cheersout_500/ra_cytoimmgen_H3K27ac_uniquePeaks.txt"), mean=mean_cytoimmgen, celllist=cytoimmgen_celllist)

# Get overlapping SNPs
cytoimmgen_H3K27ac_snps <- list()
cytoimmgen_H3K27ac_snps[["Crossdisorder psych"]] <- read.table(here("res/cheersout_500/crossdis_cytoimmgen_H3K27ac_SNPsOverlappingPeaks.txt"), sep="\t", header=T)
cytoimmgen_H3K27ac_snps[["Schizophrenia"]] <- read.table(here("res/cheersout_500/szs_cytoimmgen_H3K27ac_SNPsOverlappingPeaks.txt"), sep="\t", header=T)
cytoimmgen_H3K27ac_snps[["Alzheimer's Disease"]] <- read.table(here("res/cheersout_500/alz_cytoimmgen_H3K27ac_SNPsOverlappingPeaks.txt"), sep="\t", header=T)
cytoimmgen_H3K27ac_snps[["BMI"]] <- read.table(here("res/cheersout_500/bmi_cytoimmgen_H3K27ac_SNPsOverlappingPeaks.txt"), sep="\t", header=T)
cytoimmgen_H3K27ac_snps[["ADHD"]] <- read.table(here("res/cheersout_500/adhd_cytoimmgen_H3K27ac_SNPsOverlappingPeaks.txt"), sep="\t", header=T)
cytoimmgen_H3K27ac_snps[["Depression"]] <- read.table(here("res/cheersout_500/mdd_cytoimmgen_H3K27ac_SNPsOverlappingPeaks.txt"), sep="\t", header=T)
cytoimmgen_H3K27ac_snps[["Bipolar Disorder"]] <- read.table(here("res/cheersout_500/bip_cytoimmgen_H3K27ac_SNPsOverlappingPeaks.txt"), sep="\t", header=T)
cytoimmgen_H3K27ac_snps[["Rheumatoid arthritis"]] <- read.table(here("res/cheersout_500/ra_cytoimmgen_H3K27ac_SNPsOverlappingPeaks.txt"), sep="\t", header=T)

cytoimmgen_H3K27ac_snps <- lapply(cytoimmgen_H3K27ac_snps, function(x){x$X <- NULL; return(x)})

# Subselect SNPs overlapping the peaks of interest
pull_overlapping_snps = function(choice_peaks, snps){
  snps$peak <- paste(snps$chr, snps$start, snps$end, sep="_")
  overlaps <- sapply(choice_peaks, function(x) {snps[snps$peak %in% x,"snp"]}, USE.NAMES = TRUE)
}

overlaps <- sapply(names(cytoimmgen_H3K27ac_snps), function(x){pull_overlapping_snps(choice_peaks=rownames(peaks[[x]]), snps=cytoimmgen_H3K27ac_snps[[x]])})

```

Make heatmap of individual peaks for crossdisorder
```{r}
peaks_choice = peaks$Crossdisorder

#Heatmap of ranks
library(pheatmap)
library(RColorBrewer)
breaksList = seq(1, 2, by = 0.1)
pheatmap::pheatmap(peaks_choice, 
         annotation_col =  anno, 
         annotation_colors = annoCols[1], 
         border_color = F,
         color = colorRampPalette(brewer.pal(n = 9, name = "Blues"))(length(breaksList)),
         breaks = breaksList,
         treeheight_row = 0, treeheight_col = 0, filename=here("pics/pheatmap_cheers_cytoimmgen_enhancers_peaks_crossdis.pdf"), width=10, height=25)
# dev.off()

```

Make heatmap of individual peaks for MDD
```{r}
peaks_choice = peaks$Depression

#Heatmap of ranks
library(pheatmap)
library(RColorBrewer)
peaks_choice = peaks$Depression
breaksList = seq(1, 2, by = 0.1)
pheatmap::pheatmap(peaks_choice, 
         annotation_col =  anno, 
         annotation_colors = annoCols[1], 
         border_color = F,
         color = colorRampPalette(brewer.pal(n = 9, name = "Blues"))(length(breaksList)),
         breaks = breaksList, cluster_rows = T, cluster_cols=F,
         filename=here("pics/pheatmap_cheers_cytoimmgen_enhancers_peaks_dep.pdf"), width=10, height=17
         )

```


FIGURE: Are the T cell overlaps for MDD the same as the peak overlaps for BMI?
```{r}

print("Number of MDD peaks")
print(nrow(peaks$Depression)) # 108
print("Number of BMI peaks")
print(nrow(peaks$BMI)) # 1022

rownames(peaks$Depression) %in% rownames(peaks$BMI) %>% table() # 9 of 108 total MDD are overlaps with BMI

(repper <- dim(anno)[1])

trying <- merge(peaks$Depression, peaks$BMI, by=0, all.x=T, suffixes=c(".Dep",".BMI"))
dim(trying)
rownames(trying) <- trying$Row.names
trying$Row.names <- NULL
# Reverse within each condition so x-axis is ordered like other figures
trying <- trying[,c(seq(repper,1,-1), seq(repper*2,(repper+1),-1))]

# Make Bespoke anno without clustering
# Must rename in accordance with suffixes above
anno_dep <- anno 
rownames(anno_dep) <- paste0(rownames(anno),".Dep")
anno_bmi <- anno
rownames(anno_bmi) <- paste0(rownames(anno),".BMI")

# Get peak clustering order from internals of pheatmap using MDD alone, then use this as the fixed order for both BMI and MDD. 
# Silent pheatmap to get clustering of MDD alone
out <- pheatmap::pheatmap(peaks$Depression, 
         cluster_cols = F, silent = T)
head(out$tree_row$labels)
head(out$tree_row$order)
trying[is.na(trying)] <- 0
trying_reordered <- trying[out$tree_row$labels[out$tree_row$order],]
annoBespoke <- rbind(anno_dep, anno_bmi)
annoBespoke$condition <- c(rep("Depression",repper), rep("BMI",repper))
annoBespoke$cell_type <- NULL
peaks_choice=trying_reordered # 
breaksList = seq(1, 2, by = 0.1)
pheatmap::pheatmap(peaks_choice, 
         annotation_col =  annoBespoke, 
         annotation_colors = annoCols[1], 
         border_color = F,
         color = colorRampPalette(brewer.pal(n = 9, name = "Blues"))(length(breaksList)),
         breaks = breaksList,
         cluster_cols = F, cluster_rows = F, 
         gaps_col = dim(anno)[1],
         filename=here("pics/pheatmap_cheers_cytoimmgen_enhancers_peaks_dep_with_bmi_overlapsREV.pdf"), width=17, height=13)

```

Make Venns and upset plot
```{r}

library(ggVennDiagram)
x <- list(`Crossdisorder\npsych`=rownames(peaks$Crossdisorder), 
          Schizophrenia=rownames(peaks$Schizophrenia),
          Depression=rownames(peaks$Depression),
          `Rheumatoid\narthritis`=rownames(peaks$`Rheumatoid arthritis`)
          )
p <- ggVennDiagram(x, label_alpha=0, label="count") + guides(fill=FALSE)
p
ggsave(here("pics/CHEERS_cytoimmgen_enhancers_venn1.pdf"), p, width=7, height=4)


x <- list(`Crossdisorder\npsych`=rownames(peaks$Crossdisorder), 
          Schizophrenia=rownames(peaks$Schizophrenia),
          Depression=rownames(peaks$Depression),
          BMI=rownames(peaks$BMI)
          )
p <- ggVennDiagram(x, label_alpha=0, label="count") + guides(fill=FALSE)
p
ggsave(here("pics/CHEERS_cytoimmgen_enhancers_venn3.pdf"), p, width=7, height=4)


x <- list(`Bipolar disorder`=rownames(peaks$`Bipolar Disorder`), 
          Schizophrenia=rownames(peaks$Schizophrenia),
          Depression=rownames(peaks$Depression),
          BMI=rownames(peaks$BMI)
          )
p <- ggVennDiagram(x, label_alpha=0, label="count") + guides(fill=FALSE)
p
ggsave(here("pics/CHEERS_cytoimmgen_enhancers_venn4.pdf"), p, width=7, height=4)


x <- list(`Alzheimer's\nDisease`=rownames(peaks$`Alzheimer's Disease`),
          `Crossdisorder\npsych`=rownames(peaks$Crossdisorder), 
          Depression=rownames(peaks$Depression),
          BMI=rownames(peaks$BMI)
          )
p <- ggVennDiagram(x, label_alpha=0, label="count") + guides(fill=FALSE)
p
ggsave(here("pics/CHEERS_cytoimmgen_enhancers_venn2.pdf"), p, width=7, height=4)

x <- list(ADHD=rownames(peaks$ADHD),
          Crossdisorder=rownames(peaks$Crossdisorder), 
          Depression=rownames(peaks$Depression),
          BMI=rownames(peaks$BMI)
          )
p <- ggVennDiagram(x, label_alpha=0, label="count") + guides(fill=FALSE)
p
ggsave(here("pics/CHEERS_cytoimmgen_enhancers_venn5.pdf"), p, width=7, height=4)

library(ComplexHeatmap)
x <- list(`Crossdisorder psych`=rownames(peaks$Crossdisorder), 
          Schizophrenia=rownames(peaks$Schizophrenia),
          `Bipolar disorder`=rownames(peaks$`Bipolar Disorder`), 
          Depression=rownames(peaks$Depression),
          ADHD=rownames(peaks$ADHD),
          BMI=rownames(peaks$BMI),
          `Alzheimer's Disease`=rownames(peaks$`Alzheimer's Disease`),
          `Rheumatoid arthritis`=rownames(peaks$`Rheumatoid arthritis`)
          )
mtx <- ComplexHeatmap::make_comb_mat(x)

# FIGURE: upset plot
pdf(file=here("pics/cytoimmgen_upset_intersections.pdf"), width=10, height=6)
m <- mtx[comb_degree(mtx)>1]
ComplexHeatmap::UpSet(m, set_order = names(x), 
                      top_annotation = upset_top_annotation(m, gp = gpar(fill='white'), height = unit(7,"cm"), annotation_name_rot = 0), 
                      right_annotation = upset_right_annotation(m, gp = gpar(fill='white'), width = unit(7,"cm"), annotation_name_rot = 0))

dev.off()

```

Look at shared peaks for crossdisorder, szs and MDD
```{r}
# Union of implicated peaks across Xdiss-MDD-SZS

tmpdep = peaks$Depression[!rownames(peaks$Depression) %in% rownames(peaks$Crossdisorder),]
tmpszs = peaks$Schizophrenia[!rownames(peaks$Schizophrenia) %in% rownames(peaks$Crossdisorder),]
union <- rbind(peaks$Crossdisorder, tmpszs, tmpdep)
dim(union)

# Pheatmap, allowing to cluster, and extract the order of peaks

# Get peak clustering order from internals of pheatmap
out <- pheatmap::pheatmap(union, cluster_cols = T, silent = T)
union_reordered <- union[out$tree_row$labels[out$tree_row$order],c(seq(repper,1,-1))]

rowannot <- data.frame(
  Crossdisorder = as.factor((rownames(union_reordered) %in% rownames(peaks$Crossdisorder))),
  Schizophrenia = as.factor((rownames(union_reordered) %in% rownames(peaks$Schizophrenia))),
  Depression = as.factor((rownames(union_reordered) %in% rownames(peaks$Depression))),
  row.names = rownames(union_reordered)
)
head(rowannot)

annoCols = list(cell_time = timepointlabeller,
                Schizophrenia=c("TRUE"="dark red","FALSE"="light grey"), Crossdisorder=c("TRUE"="dark red","FALSE"="light grey"), Depression=c("TRUE"="dark red","FALSE"="light grey"))

peaks_choice = union_reordered

# View to see all peaks
library(pheatmap)
library(RColorBrewer)
breaksList = seq(1, 2, by = 0.1)
pheatmap::pheatmap(peaks_choice, 
         annotation_col =  anno, 
         annotation_colors = annoCols, 
         annotation_names_row = T,
         annotation_row = rowannot,
         border_color = F,
         color = colorRampPalette(brewer.pal(n = 9, name = "Blues"))(length(breaksList)),
         breaks = breaksList,
         cluster_rows = F, cluster_cols = F,
         filename=here("pics/pheatmap_cheers_cytoimmgen_enhancers_union_rowannot.pdf"), width=10, height=40)

```

T-cell late activation peaks selected from above plot
```{r}
clump_union_late_cd4_activation1 <- c("chr10_102860759_102861039",
"chr17_18044981_18045390",
"chr9_93479394_93555145",
"chr9_93443391_93478580",
"chr12_110277872_110355989",
"chr5_138505654_138578588",
"chr14_103555078_103571667",
"chr1_173818608_173832564",
"chr2_199909685_199917329",
"chr8_38161192_38200419",
"chr14_102762098_102842127",
"chr14_103651138_103659209",
"chr1_51839557_51840042",
"chr11_46455125_46458833",
"chr2_27806852_27812367")

clump_union_late_cd4_activation2 <- c("chr11_46513514_46516633",
"chr11_46574378_46577301",
"chr14_103572314_103580026",
"chr3_52829680_52854555",
"chr11_46419164_46422656",
"chr12_118385640_118394696",
"chr11_46539665_46549623",
"chr5_138385021_138411450",
"chr5_154267126_154279634",
"chr11_46428338_46433483",
"chr11_46435152_46437412",
"chr3_49849607_49850058",
"chr5_154261006_154261449",
"chr11_46399922_46403048",
"chr11_46406942_46417533",
"chr17_18055388_18062637",
"chr5_154226463_154237900",
"chr3_49826744_49848468",
"chr11_46471560_46486285",
"chr20_46059114_46067543",
"chr1_243389226_243392351",
"chr11_46490479_46512079",
"chr5_62197263_62202925")

clump_union_late_cd4_activation3 <- c("chr2_207181680_207185557",
"chr2_207218442_207261835",
"chr5_61302458_61309191",
"chr13_31275256_31275785",
"chr1_51843127_51845502",
"chr15_90735800_90888696",
"chr5_154289840_154303036",
"chr5_88246379_88256745",
"chr5_62221766_62243169",
"chr2_200370276_200393953",
"chr15_84530565_84536387",
"chr6_28396163_28401899",
"chr6_25236231_25237674",
"chr1_66617270_66628541",
"chr17_17987580_17993504",
"chr17_18085663_18111161",
"chr14_103687819_103726312",
"chr14_103731996_103743690",
"chr2_232726821_232731946",
"chr2_199935762_199964974",
"chr22_41204309_41231646",
"chr19_19311021_19339648",
"chr6_27877340_27880978",
"chr6_27804709_27840700",
"chr6_27861293_27876636")

clump_union_late_cd4_activation_all <- c(clump_union_late_cd4_activation1, clump_union_late_cd4_activation2, clump_union_late_cd4_activation3)

length(clump_union_late_cd4_activation_all) 
anno$cell_type <- NULL

# FIGURE Pheatmap focused on late cd4 activation
peaks_choice = union_reordered[rownames(union_reordered) %in% clump_union_late_cd4_activation_all,]
pheatmap::pheatmap(peaks_choice, 
         annotation_col =  anno, 
         annotation_colors = annoCols, 
         annotation_names_row = T,
         annotation_row = rowannot,
         border_color = F,
         color = colorRampPalette(brewer.pal(n = 9, name = "Blues"))(length(breaksList)),
         breaks = breaksList,
         cluster_rows = F, cluster_cols = F,
         filename=here("pics/pheatmap_cheers_cytoimmgen_enhancers_union_rowannot_cd4_late.pdf"), width=8.3, height=9.7)
```

```{r}
save(cytoimmgen_celllist, cytoimmgen_H3K27ac_snps, cytoimmgen_H3K27ac_p, peaks, overlaps, anno, annoCols, rowannot, union_reordered, clump_union_late_cd4_activation_all, file = here("res/cheers_cytoimmgen_processed.RData"))
```
